---
title: "Cluster Analysis for Crime Data"
author: "Clustering INSOFE Lab Assignment"
date: "05 Nov 2017"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
---

**NOTE** Before starting this assignment please remember to clear your environment, you can do that by running the following code chunk

```{r}


```

# Agenda

* Get the data

* Data pre-processing

* Explore the data

* Hierarchical Clustering

* Kmeans Clustering

* Visualising Clusters and Evaluation


# Problem Description

* In the following Unsupervised Learning activity, you will perform cluster analysis on a dataset that has arrests per 100,000 residents for assault, murder, and rape in each of the 50 US states

* The variable names in the dataset are self explanatory

* So, you will cluster US states based on the crime rates, which can then be passed on to public policy experts to gain insight from it


# Reading & Understanding the Data

* Read in the dataset

```{r}

# Use the setwd() function to get to the directory where the data is present



```

* Use the str() and summary() functions to get a feel for the dataset.

```{r}



```

* Take a look at the data using the "head()" and "tail()" functions

```{r}



```

# Data pre-processing

* Check for the number of missing values in the dataset

```{r}




```


* Convert the State names into row names and remove that variable from the dataset

```{r}





```

* Standardize and scale the data

```{r}



```

# Data exploration

* Visualise the distances between the individual observations using the fviz_dist()

```{r, fig.width=12, fig.height=8}








```

# Hierarchical Clustering

* Cluster the data using the Ward algorithm

```{r}






```

* Plot the dendogram for hierarchical clustering

```{r, fig.height=5, fig.width=10}



```

* Cut the tree to 4 clusters

```{r}









```

* Plot a new dendogram, with each of the clusters being surrounded by a border, using the rect.hclust() function

```{r, fig.height=5, fig.width=10}





```


# K-Means Clustering

* Build a basic kmeans model with k = 2

```{r}





```

* Build a scree plot and choose a "k"

```{r}










```

* Choose a k and cluster the data

```{r}

# We can choose a k = 4, 5 or 6



```

* Visualise the clusters by plotting the data points on the first two principal components

```{r, fig.height=5, fig.width=8}



```

